### 2015-07-12 narrative/intro.tex
rewrote the perspective portions

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline

\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

To accomplish this the application and storage teams developed new
application specific data reduction techniques and added them to the ADIOS
I/O middleware layer. A rudimentary discovery system was also created to
track what data was on disk and what was on the tape archive. Initially this
solution was sufficient, but eventually a new problem was introduced:
Although the team had reserved resources for the run, the entire file system
was shared \cite{shipman2012next} resulting in very high I/O variability
\cite{lofstead:2010:io-variability}. Addressing this variability, caused by
contention on shared resources, is an ongoing concern.

@@

@@

\paragraph{Simulation Perspective}
\label{subsec:sim-perspective}
A simulation scientist should ideally be able to
obtain an estimate of how much time will be required to write data,
and how much storage space is available during the simulations lifetime.
Further, they would like Quality of Service (QoS) assurances, and that
issues such as varying bandwidth can be handled appropriately and 
automatically. For example, a set of user-specified rules would allow 
the system to make autonomic decisions at runtime
~\S~\ref{sec:managing-data-life}~\S~\ref{sec:qos}.
%

Central to our approach is giving users the ability to define the utility of their data    
 \S\ref{sec:data-description}.
This includes the ability to assign varying levels of importance to individual data objects in their entirety, or to specific
subsets and pieces.
Key research challenges include: 1) developing an expressive and easy to use a program interface, 2) formulating a system model that can support 
autonomic decision, for example, for saving data based on user-defined importance and utility of data~\S~\ref{sec:managing-data-life}),
, and 3) defining the interface between the system and the back-end storage for both single and multiple running applications~\S~\ref{sec:qos}.

In the scenario above, the scientist would then specify a set of rules that would guide I/O and data management 
at runtime. For example, if system resources are available and the estimated write times acceptable, the write 
can proceed, but if this is not the case, the write may be postponed to a later time-step, or the data reduced using
an in-situ techniques before it is saved. Data ``importance or utility'' can be used to place various data at  
proper SSIO layers -- data with the highest utility (e.g., based on user defined importance or the frequency of access) 
is maintained in faster storage for long periods of time, while data with lower utility can be migrated to lower layers 
in the storage hierarchy.

@@

\paragraph{Storage System Perspective}
\label{subsec:storage-perspective}
From the perspective of the storage system, it is necessary to manage the data from
a given user request amongst all of the requests from all users, and try to optimize the
entire system accordingly. Today the problem is that when a user writes there is a
tremendous amount of interference from other users on the system who may be
writing or reading at the same time, leading to unpredictable performance.
By requiring a certain QoS~\S\ref{sec:qos}, the system could potentially lock out and throttle other users. 
The storage system must also automatically move data from one tier of the storage to
another, without affecting the QoS that was offered at request
time. The storage system must be able to organize
data amongst all of the tiers of storage in a known, consistent way.

One of our research objectives is to understand how the
storage system can interact with the middleware   so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. We need to maintain 
QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
the different layers to inform any migration/eviction decisions, and it
is also critical that the storage system can manage (discover and retrieve) the data across all of the
layers automatically, without user intervention. 
Central to our approach is that the storage system needs to define bounds of time estimates
(min time, max time), and as our system evolves, we want to allow plug-ins which can help
reduce the range of the bounds.
Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
~\S\ref{sec:naming-discovery}.

@@

\paragraph{Reading  Perspective}
\label{subsec:reading-perspective}
From the perspective of reading, this project will look at key I/O patterns such as checkpoint-restart (CR), in-situ analytics and post-processing. 
For example, in case of CR, we will support both, traditional file-based approaches leveraging fast storage tiers
such as NVRAM (non-volatile RAM), as well as memory-based approach based on retrieving
data directly from NVRAM. In case of in-situ and post-processing,  an important aspect is the latency -- here reducing latency and providing QoS 
 becomes important. 
 When someone is doing interactive analysis and expects that the data they read in is about
10s, but it turns into 1,000s, the user typically tends to either wait
another day for doing their analysis or gets frustrated and reads much less
data than they really need to. Furthermore the user tends to read in just a
small sample of the data without much knowledge if they are missing
important pieces of the data~\S\ref{sec:migration}.

>>>>

@@

To accomplish this, the application and storage teams developed new
application specific data reduction techniques and added them to the ADIOS
I/O middleware layer. A rudimentary discovery system was also created to
track what data was on disk and what was on the tape archive. Initially this
solution was sufficient, but eventually a new problem was introduced:
Although the team had reserved resources for the run, the entire file system
was shared \cite{shipman2012next} resulting in very high I/O variability
\cite{lofstead:2010:io-variability}. Addressing this variability, caused by
contention on shared resources, is an ongoing concern.

@@

\begin{table}[htbp]
\small
\vspace{2ex}
\begin{center}
\caption {Relevance to SSIO}
\label{tab:call}
\begin{tabular}{ | p {0.2cm} | p{5cm} | p{7.5cm} | p{2.2cm} |}\hline
\multicolumn{2}{|c|}{\bf Call area} & \multicolumn{1}{|c|}{\bf Relevant S2E2 work} & \multicolumn{1}{|c|}{\bf Proposed research} \\\hline\hline
1 & Data abstractions & Create new APIs  and annotations to guide data placement across the storage system & \S\ref{sec:data-description} \\\hline
2 & Mapping science data models onto hierarchical storage &  Enable science-driven mapping, re-organization, reduction and prioritization of data chunks derive from application objects & \S\ref{sec:data-refactor}, \S\ref{sec:data-description},  \S\ref{sec:managing-data-life}\\\hline
3 & Mechanisms for data movement across the storage hierarchy & Enable science-driven data placement and migration based on policies& \S \ref{sec:managing-data-life},  \S\ref{sec:migration} \\\hline
4 &  Extend the interaction of middleware with the storage system & Manage data placement and migration between the application, middleware and storage system using an autonomic cross-layer approach & \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery}  \\\hline
5 & Design and implement of I/O middleware architectures  & Create APIs to guide the placement of chunks of data across the SSIO and autonomically move data& \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:naming-discovery} \\ \hline
6 &Approaches to improve the ability of SSIO software to support checkpoint/restart &  We will have utility functions defined for the lifetime of data, and allow CR data to be automatically placed in the fastest storage and then later purged & \S\ref{sec:data-description}, \S\ref{sec:managing-data-life}, \S\ref{sec:qos}, \S\ref{sec:data-description} \\ \hline
\end{tabular}
\end{center}
\vskip -0.5cm
\end{table}

@@

\paragraph{Simulation Perspective}
\label{subsec:sim-perspective}
A simulation scientist should ideally be able to
obtain an estimate of how much time will be required to write data,
and how much storage space is available during the simulations lifetime.
Further, they would like Quality of Service (QoS) assurances, and that
issues such as varying bandwidth can be handled appropriately and 
automatically. For example, a set of user-specified rules would allow 
the system to make autonomic decisions at runtime
~\S~\ref{sec:managing-data-life}~\S~\ref{sec:qos}.
%
Our approach is based on giving users the ability to define the utility of
their data \S\ref{sec:data-description}. This includes the ability to assign
varying levels of importance to individual data objects in their entirety,
or to specific subsets and pieces. Key research challenges include: 1)
developing an expressive and easy to use a program interface, 2) formulating
a system model that can support autonomic decision, for example, for saving
data based on user-defined importance and utility of
data~\S~\ref{sec:managing-data-life}), and 3) defining the interface between
the system and the back-end storage for both single and multiple running
applications~\S~\ref{sec:qos}.

The scientist will be able to specify a set of rules that
would guide I/O and data management at runtime. For example, if system
resources are available and the estimated write times acceptable, the write
can proceed, but if this is not the case, the write may be postponed to a
later time-step, or the data reduced using an in-situ techniques before it
is saved. Data ``importance or utility'' can be used to place various data
at proper SSIO layers -- data with the highest utility (e.g., based on user
defined importance or the frequency of access) is maintained in faster
storage for long periods of time, while data with lower utility can be
migrated to lower layers in the storage hierarchy.

@@

\paragraph{Storage System Perspective}
\label{subsec:storage-perspective}
From the perspective of the storage system, it is necessary to manage a
single I/O request from a user in the context of all requests from all
users, and try to optimize the entire system accordingly. This entails
providing quality of service (~\S\ref{sec:qos} to the user to minimize
resource contention, distribute the data across all of the tiers of storage
in a known consistent manner (\S~\ref{sec:init-plac-data}), and automatically move data from one part of
the storage hierarchy to another (~\ref{sec:migration}).

One of our research objectives is to understand how the
storage system can interact with the middleware   so
that the data with the highest utility is kept on the fastest layers of the
storage system for as long as specified. We need to maintain 
QoS and therefore, be able to provide fuzzy estimates of the time it will take to write/read data from
the different layers to inform any migration/eviction decisions, and it
is also critical that the storage system can manage (discover and retrieve) the data across all of the
layers automatically, without user intervention. 
Our approach here is that the storage system needs to define bounds of time estimates
(min time, max time), and as our system evolves, we want to allow plug-ins which can help
reduce the range of the bounds.
Overall, our goal is to enable predictable performance acorss the storage system, and we will develop mechanisms to enable this
~\S\ref{sec:naming-discovery}.

@@

\paragraph{Reading  Perspective}
\label{subsec:reading-perspective}

From the reading perspective, the data must be accessible in a fast and
predictable manner for all read access patterns, with different access
patterns having their own requirements for latency, throughput and
consistency. Here the user must be able to describe the read patterns and
query for performance estimations (~\S~\ref{sec:data-description}), be able
to quickly access the most important data (\S~\ref{sec:data-refactor}) and recieve
performance gurantees (~\S\ref{sec:qos}) on the time required to complete
the read request. The challenge here is to address the variety of read
patterns that drive knowledge discover.

For example, consider the case of Checkpoint/Restart (C/R) which needs a lot
of data at high throughput to complete the restart in case of application
failure, with latency being less important due to the batch nature of the
request, and the data must be of sufficient quality to support an
application restart without significant errors. 
This is in contrast to interactive analysis, where the most important metric
is the latency to start the analysis process. The quality here is more
variable and based greatly on the expectations of the user, but performance
predictability is important due to the interactive nature of the process.

Our research objective here is to explore the interface that reading code
can utilize to discover the important properites of data, the expectations
about the time to access the data, and make decisions on the quality of data
required to complete the task.
